---
title: "Peekbank image import"
author: "Tarun Sepuri"
date: "`r format(Sys.time(), '%d %B, %Y')`"
format:
  html:
    toc: true
    toc_depth: 3
    toc_float: true
editor_options: 
  chunk_output_type: inline
---
# Download peekbankr

If you haven't downloaded peekbankr (https://github.com/langcog/peekbankr) yet, be sure to do so first by uncommenting the lines below.

```{r}
knitr::opts_chunk$set(cache = FALSE, arn = FALSE,warning=FALSE, message = FALSE,cache.lazy = FALSE)
```

# Preliminaries and data loading

Load packages. Since it takes a while to download and join the data, you probably want to just do that once, and then save the resulting dataset. Setting the parameter FIRST_TIME to FALSE after you run the script the first time allows you to bypass the data download process on subsequent runs. You can also use the most recent data file uploaded to GitHub.

```{r}
FIRST_TIME = FALSE # set to true first time to download data from DB

if (FIRST_TIME) {
  install.packages("remotes") # can also use devtools
  remotes::install_github("peekbank/peekbankr", force=TRUE)
}

library(peekbankr)
library(tidyverse)
library(here)
library(RColorBrewer)
library(cowplot)
library(magrittr)
theme_set(theme_cowplot())
source("helpers.R")
```

```{r}
if (FIRST_TIME) {
  
  conn = connect_to_peekbank("2026.1",port=3307)
  administrations <- peekbankr::get_administrations(connection = conn) %>% dplyr::collect()
  subjects <- peekbankr::get_subjects(connection = conn) %>% dplyr::collect()
  aoi_timepoints <- peekbankr::get_aoi_timepoints(connection = conn) %>% dplyr::collect()
  
  # need to reconnect since aoi_timepoints takes a long time and causes our connection to drop
  conn <- connect_to_peekbank("2026.1", port=3307)
  all_stimuli <- peekbankr::get_stimuli(connection = conn) %>% dplyr::collect()
  trial_types <- peekbankr::get_trial_types(connection = conn) %>% dplyr::collect()
  trials <- peekbankr::get_trials(connection = conn) %>% dplyr::collect()
  saveRDS(administrations, here("intermediates/administrations.Rds"))
  saveRDS(subjects, here("intermediates/subjects.Rds"))
  saveRDS(aoi_timepoints, here("intermediates/aoi_timepoints.Rds"))
  saveRDS(all_stimuli, here("intermediates/all_stimuli.Rds"))
  saveRDS(trial_types, here("intermediates/trial_types.Rds"))
  saveRDS(trials, here("intermediates/trials.Rds"))
  DBI::dbDisconnect(conn)
} else {
  administrations <- readRDS(here("intermediates/administrations.Rds"))
  subjects <- readRDS(here("intermediates/subjects.Rds"))
  aoi_timepoints <- readRDS(here("intermediates/aoi_timepoints.Rds"))
  all_stimuli <- readRDS(here("intermediates/all_stimuli.Rds"))
  trial_types <- readRDS(here("intermediates/trial_types.Rds"))
  trials <- readRDS(here("intermediates/trials.Rds"))
}
```



dataset level subject counts - more on this in the Descriptives section
```{r}
subj_counts <- administrations |> left_join(subjects) |>
  distinct(subject_id, .keep_all=TRUE) |>
  summarize(n = n(), .by=c("dataset_name", "dataset_id"))
```

# Downloading stimuli files

```{r}
if (FIRST_TIME) {
  conn = connect_to_peekbank("2026.1",port=3307)
  downloaded_stimuli <- peekbankr::download_stimuli(conn, local_base_dir = "stimulus_data")
  DBI::dbDisconnect(conn)
  saveRDS(downloaded_stimuli, here("intermediates/downloaded_stimuli.Rds"))
} else {
  downloaded_stimuli <- readRDS(here("intermediates/downloaded_stimuli.Rds"))
}
stimuli_dataset_information <- downloaded_stimuli |>
  group_by(dataset_id, dataset_name) |>
  summarise(
    stimuli_with_path = n(),  # Total stimuli with defined paths in this dataset
    total_stimuli = length(all_stimuli$stimulus_id[all_stimuli$dataset_id == dataset_id]),
    prop_with_defined_path = (stimuli_with_path / total_stimuli),  # Percentage of total stimuli
  )

stimuli_dataset_information

```


Just plotting dataset stimuli information
```{r}
ggplot(stimuli_dataset_information, aes(x=prop_with_defined_path, y=stimuli_with_path, color=dataset_name)) +
  xlab("Proportion of stimuli available") +
  ylab("Number of stimuli available") +
  geom_point(size = 4, alpha=0.3) 
```

Create a CSV file of all of the pairs of images in all of the trials that we have stimuli files for
```{r}
stimuli_trial_info <-  downloaded_stimuli |> 
      select(stimulus_id, image_description, local_stimulus_path, english_stimulus_label, original_stimulus_label) |>
      mutate(english_stimulus_label = str_to_lower(english_stimulus_label), original_stimulus_label = str_to_lower(original_stimulus_label))

downloaded_stimuli_trials <- trial_types |>
  left_join(
    stimuli_trial_info |> 
      rename_with(~ paste0(.x, "_target"), everything()),
    by = c("target_id" = "stimulus_id_target")
  ) |>
  left_join(
    stimuli_trial_info |> 
      rename_with(~ paste0(.x, "_distractor"), everything()), 
    by = c("distractor_id" = "stimulus_id_distractor")
  ) |>
  # This ensures that both the target and distractor do exist.
  filter(!is.na(local_stimulus_path_distractor) & !is.na(local_stimulus_path_target)) |>
  mutate(
    unique_pair = pmap_chr(list(target_id, distractor_id), ~ paste(sort(c(.x, .y)), collapse = "_"))
  ) |>
  select(-condition, -lab_trial_id, -aoi_region_set_id, -point_of_disambiguation)

# unique pairs keep trial-distractor pairing, we only need to find similarities for the trials that we but we will probably need to do pair-wise comparisons down the line
unique_stimuli_pairs <- downloaded_stimuli_trials |>
  rename(text1=english_stimulus_label_target, text2=english_stimulus_label_distractor, text1_description=image_description_target, text2_description=image_description_distractor, image1=local_stimulus_path_target, image2=local_stimulus_path_distractor,dataset_name=dataset_name) |> distinct(unique_pair, image1, image2, text1, .keep_all=TRUE) 

write.csv(unique_stimuli_pairs,file=here("data/peekbank_stimuli.csv"))
```

```{r}
paste("Datasets")
unique(downloaded_stimuli_trials$dataset_name)
```

# Processing stimuli AOIs
```{r}
downloaded_aois <- trials |>
  filter(excluded == 0 & trial_type_id %in% downloaded_stimuli_trials$trial_type_id) |>
  left_join(aoi_timepoints) |>
  mutate(accuracy = ifelse(aoi == "target", 1, 0),
    not_looking_away = aoi == "target" | aoi == "distractor",
    accuracy = ifelse(not_looking_away, accuracy, NA))

saveRDS(downloaded_aois, here("intermediates/downloaded_aois.Rds"))
```

```{r}
count(trials |> left_join(trial_types) |>
  distinct(target_id, distractor_id))
```

## Summarize usable trials with stimuli
helpers
```{r}
# Function to summarize whether a trial is usable based on whether the subject is looking at the screen for greater than 50% of the critical window
summarize_subj_usable_trials <- function(data, critical_window, suffix, additional_fields=NULL) {
  additional_fields <- additional_fields %||% list()
  
  data %>%
    filter(t_norm >= critical_window[1] &
             t_norm <= critical_window[2]) %>%
    group_by(administration_id, trial_id, trial_type_id) %>%
    summarize(
      length = n(),
      usable_frames = sum(not_looking_away, na.rm = TRUE),
      percent_usable = usable_frames / length,
      usable = ifelse(percent_usable >= 0.5, 1, 0), # usable if at least 50% looking
      mean_target_looking = mean(accuracy, na.rm = TRUE),
      !!!additional_fields,
    ) %>%
    rename_with(~ paste0(., "_", suffix), -c(administration_id, trial_id, trial_type_id))
}

# Function to compute whether a trial is usable based on whether both the critical window and the baseline window are usable
compute_usable_trial <- function(baseline_col, critical_col) {
  case_when(
    is.na(baseline_col) ~ 0,
    is.na(critical_col) ~ 0,
    baseline_col == 1 & critical_col == 1 ~ 1,
    TRUE ~ 0
  )
}

# Calculate mean, standard deviation, standard error and confidence intervals for data grouped across two variables
summarized_data <- function(data, x_var, y_var, group_var) {
  return(data |>
           group_by(across(all_of(c(x_var, group_var)))) |>
           summarize(
                   #across(everything(), ~ if (n_distinct(.) == 1) first(.) else NA),
                    mean_value = mean(.data[[y_var]], na.rm = TRUE),
                     sd_value = sd(.data[[y_var]], na.rm = TRUE),
                     N = n(),
                     se = sd_value / sqrt(n()),
                     ci=qt(0.975, N-1)*sd_value/sqrt(N),
                     lower_ci=mean_value-ci,
                     upper_ci=mean_value+ci,
                     .groups = 'drop') |>
           select(where(~ !all(is.na(.))))
  )
}
```

using same settings as visvocab
```{r}
critical_window <- c(300,3500)
critical_window_short <- c(300,1800)
baseline_window <- c(-2000,0)

# summarize critical window
summarize_subj_usable_trials_critical_window <- summarize_subj_usable_trials(
  data = downloaded_aois,
  critical_window = critical_window,
  suffix = "critical_window"
)
# summarize short critical window
summarize_subj_usable_trials_critical_window_short <- summarize_subj_usable_trials(
  data = downloaded_aois,
  critical_window = critical_window_short,
  suffix = "critical_window_short"
)

# combine into one dataset
summarize_subj_usable_trials_critical_window <- summarize_subj_usable_trials_critical_window %>%
  left_join(summarize_subj_usable_trials_critical_window_short)

# summarize baseline window information
summarize_subj_usable_trials_baseline_window <- summarize_subj_usable_trials(
  data = downloaded_aois,
  critical_window = baseline_window,
  suffix = "baseline_window")

min_baseline_values <- downloaded_aois |>
  group_by(administration_id, trial_id, trial_type_id) |>
  filter(t_norm == min(t_norm)) |>
  rename(min_time = t_norm) |>
  distinct()

#overall usable trials
summarize_subj_trials <- downloaded_aois %>%
  distinct(administration_id, trial_id, trial_type_id) %>%
  left_join(summarize_subj_usable_trials_critical_window) %>%
  left_join(summarize_subj_usable_trials_baseline_window) %>%
  # joining in min baseline window value, for baseline window related stats we need to filter to baseline windows with -500ms or lower min value.
  left_join(min_baseline_values) %>%
  mutate(
    usable_window = compute_usable_trial(usable_baseline_window, usable_critical_window),
    usable_window_short = compute_usable_trial(usable_baseline_window, usable_critical_window_short),
    corrected_target_looking = mean_target_looking_critical_window - mean_target_looking_baseline_window,
    corrected_target_looking_short = mean_target_looking_critical_window_short - mean_target_looking_baseline_window
  )
```

```{r}
usable_trials_summarized <- summarize_subj_trials %>%
  # filtering to trials where participant looks for more than 50% of the baseline and critical windows.
  filter(usable_window == 1) |> left_join(administrations |> select(administration_id, subject_id)) |>
  # excluding participants with less than 5 trials
  group_by(subject_id) |>
  filter(n() >= 5) |>
  ungroup()
```

### Overall timecourse plot of proportion target looking 
```{r}
#summarizing within subject for each time point
summarize_subj_aois <- summarized_data(downloaded_aois |> filter(!is.na(accuracy) & trial_id %in% usable_trials_summarized$trial_id), "t_norm", "accuracy", "administration_id") |> rename(mean_accuracy = mean_value)

#summarizing across subjects for each time point
summarize_across_subj_aois <- summarized_data(summarize_subj_aois, "t_norm", "mean_accuracy", "t_norm") |> rename(accuracy = mean_value)
saveRDS(summarize_across_subj_aois, here("intermediates/summarize_across_subj_aois.Rds"))

looking_times <- ggplot(summarize_across_subj_aois,aes(t_norm,accuracy))+
  xlim(-2000,4000)+
  geom_errorbar(aes(ymin=accuracy-ci,ymax=accuracy+ci),width=0, alpha=0.2)+
  #geom_point(alpha=0.2)+
    geom_smooth(method="gam")+
  geom_vline(xintercept=0,size=1.5)+
  geom_hline(yintercept=0.5,size=1.2,linetype="dashed")+
  geom_vline(xintercept=300,linetype="dotted")+
  ylim(0,1)+
  xlab("Time (normalized to target word onset) in ms")+
  ylab("Proportion Target Looking")
looking_times
ggsave(here("figures","prop_looking_across_time.png"),looking_times,width=9,height=6,bg = "white")
```

Age-related effects
```{r}
summarize_subjects <- summarized_data(usable_trials_summarized |> 
                                        left_join(administrations), "administration_id", "corrected_target_looking", "age")
ggplot(summarize_subjects, aes(x = age, y = mean_value)) +
  geom_hline(yintercept = 0.0, linetype = "dashed") +
  geom_point(size = 4, alpha = 0.1,
             position = position_jitter(width = 0.1, seed = 123)) +
  geom_smooth(method="lm") +
  scale_x_continuous(breaks = seq(5, 75, by = 10)) +
  labs(x = "Age in months", y = "Corrected prop. of target looking by age") +
  ggtitle("Corrected proportion of target looking by age") +
  theme_minimal() +
  ggpubr::stat_cor(method = "pearson")
```


#Loading CLIP and other similarities
```{r}
similarities <- read.csv(here("data", "similarities_data.csv"))

pair_key <- function(a, b) paste(pmin(a, b), pmax(a, b), sep = "_")

# --- canonicalize the similarity tables, one per space, keyed by what it depends on ---
image_sims <- similarities |>
  transmute(img_key = pair_key(image1, image2), image_similarity) |>
  distinct()

text_sims <- similarities |>
  transmute(txt_key = pair_key(text1, text2), text_similarity, ooo_similarity) |>
  distinct()

# multimodal is anchored to the TARGET text and image (text1, image1), so can't sort
mm_sims <- similarities |>
  transmute(mm_key = paste(text1, image1, image2, sep = "_"),
            multimodal_similarity) |>
  distinct()

# --- build the trial table with the same keys, then join ---
usable_trials_imagepair_sims <- usable_trials_summarized |>
  left_join(
    downloaded_stimuli_trials |>
      transmute(trial_type_id,
                unique_pair,
                target_id,
                full_phrase_language,
                distractor_id,
                vanilla_trial,
                original_target_label = original_stimulus_label_target,
                image1 = local_stimulus_path_target,
                image2 = local_stimulus_path_distractor,
                text1  = english_stimulus_label_target,
                text2  = english_stimulus_label_distractor),
    by = "trial_type_id"
  ) |>
  mutate(
    img_key = pair_key(image1, image2),
    txt_key = pair_key(text1, text2),
    mm_key = paste(text1, image1, image2, sep = "_"),
  ) |>
  left_join(image_sims, by = "img_key") |>
  left_join(text_sims,  by = "txt_key") |>
  left_join(mm_sims,    by = "mm_key") |>
  left_join(administrations |> select(administration_id, age), by = "administration_id") |>
  mutate(trial_key = paste(target_id, distractor_id, sep="_"))
```

Filtering out trials which did not have any usable trials
```{r}
unusable_trials <- usable_trials_imagepair_sims |> filter(is.na(image_similarity)) |> distinct(unique_pair, .keep_all=TRUE) |> select(dataset_name, unique_pair, image1, image2, text1, text2, dataset_id)
unusable_trials
```

also filtering out non-vanilla trials and datasets
```{r}
usable_trials_summarized_with_sims <- usable_trials_imagepair_sims |> filter(!is.na(image_similarity))
# moving vanilla filtering to later in processing so that exploratory analyses are possible
non_vanilla_trials <- usable_trials_summarized_with_sims |> filter(vanilla_trial == 0)
vanilla_trials <- usable_trials_summarized_with_sims |> filter(vanilla_trial == 1)
non_vanilla_datasets <- non_vanilla_trials |> distinct(dataset_name) |> anti_join(vanilla_trials |> distinct(dataset_name))
non_vanilla_datasets
```

Exporting trial-level data
```{r}
write.csv(usable_trials_summarized_with_sims, here("data/usable_trials_with_similarities.csv"))
```

# Printing data counts
```{r}
things_categories <- read.csv(here("data/things_categories.csv"))
things_coverage <- usable_trials_summarized_with_sims |> semi_join(things_categories, by=c("text1"="Word")) |> semi_join(things_categories, by=c("text2"="Word"))
# trial types for which all of the trials were marked as unusable
clip_data_summarized <- summarize_similarity_data_collapsed(usable_trials_summarized_with_sims, extra_fields = c("dataset_name", "vanilla_trial"))
non_usable_unique_pairs <- similarities |>
  anti_join(usable_trials_summarized |>
  left_join(downloaded_stimuli_trials |> select(trial_type_id, unique_pair, local_stimulus_path_target, local_stimulus_path_distractor), by=c("trial_type_id")), by = c("unique_pair")) |>
  distinct(unique_pair, .keep_all = TRUE)

trial_types <- trial_types |>
 mutate(unique_pair = pmap_chr(list(target_id, distractor_id), ~ paste(sort(c(.x, .y)), collapse = "_")))

print(paste("Number of datasets:", count(usable_trials_summarized_with_sims |> distinct(dataset_name))))
print(paste("Total subjects:", count(usable_trials_summarized_with_sims|> left_join(administrations |> select(administration_id, subject_id)) |> distinct(subject_id))))
print(paste("Total trials:", count(usable_trials_summarized_with_sims)))
print(paste("Total trials covered by THINGS initially:", count(things_coverage)))
print(paste("Total trials covered by THINGS afer replacements:", count(usable_trials_summarized_with_sims |> filter(!is.na(ooo_similarity)))))

print(paste("Total target labels:", count(usable_trials_summarized_with_sims |> distinct(text1))))

cat("\nUnique label-target-distractor pairs\n")
print(paste("Number of unique label-target-distractor pairs:", count(usable_trials_summarized_with_sims |> distinct(text1, text2, dataset_name))))
print(paste("Number of unique label-target-distractor pairs covered by THINGS:", count(usable_trials_summarized_with_sims |> filter(!is.na(ooo_similarity)) |> distinct(text1, text2, dataset_name))))
print(paste("Number of unique label-target-distractor pairs with unique image-target distractor pairs:", count(usable_trials_summarized_with_sims |> distinct(text1, text2, image1, image2, dataset_name, full_phrase_language))))
print(paste("Number of unique label-target-distractor ids:", count(usable_trials_summarized_with_sims |> distinct(target_id, distractor_id, dataset_name))))
print(paste("Total label-target-distractor pairs in Peekbank:", count(trial_types |> distinct(target_id, distractor_id))))

cat("\nUnique stimulus pairs\n")
print(paste("Number of unique image and text pairs:", count(usable_trials_summarized_with_sims |> distinct(txt_key, img_key, full_phrase_language))))
print(paste("Number of unique stimulus id pairs:", count(usable_trials_summarized_with_sims |> distinct(unique_pair))))
print(paste("Number of unique image pairs with no usable trials:", count(non_usable_unique_pairs)))
print(paste("Total unique stimulus id pairs in Peekbank:", count(trial_types |> distinct(unique_pair))))

cat("\nTrial types\n")
print(paste("Number of trial types:", count(usable_trials_summarized_with_sims |> distinct(trial_type_id))))
print(paste("Total trial types in Peekbank:", count(trial_types |> distinct(trial_type_id))))
```

# Descriptive plots
### label-target-distractor combinations 
```{r}
trial_counts <- clip_data_summarized |>
  #filter(N > 10) |>
  distinct(text1, text2, dataset_name) |>
  summarize(count = n(), .by="dataset_name") |>
  arrange(count)

# Create stacked bar chart
ggplot(trial_counts, aes(x = reorder(dataset_name, -count), y = count)) +
  geom_col(fill="skyblue") +
  labs(title = "Number of unique label-target-distractor combinations",
       x = "Dataset Name", 
       y = "Count",
       fill = "Trial Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### raw trial counts
```{r}
# Function to calculate Pearson's correlation per epoch
round_to_nearest <- function(x, round_to=3) {
  round(x / round_to) * round_to
}

# rounding each participant to the closest 5
age_based_trials <- usable_trials_summarized_with_sims |> mutate(
  rounded_age = round_to_nearest(age, round_to=5)
)

clip_age_stacked_bar <- age_based_trials |>
  #filter(N > 10) |>
  group_by(rounded_age, dataset_name) |>
  summarize(trial_count = n(), .groups = "drop")

ggplot(clip_age_stacked_bar, aes(x = factor(rounded_age), y = trial_count, fill = dataset_name)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Number of Trials by Dataset and Age",
    x = "Age in months (rounded to nearest 5 month)",
    y = "Number of Trials",
    fill = "Dataset"
  ) +
  theme_minimal()
```
### Raw participant counts
#### Per dataset
```{r}
trials_with_subject_info <- usable_trials_summarized_with_sims |> left_join(administrations |> select(administration_id, subject_id)) |>
  left_join(subjects |> select(subject_id, subject_aux_data)) 

subject_counts <- trials_with_subject_info |>
  mutate(has_cdi = grepl("cdi", subject_aux_data)) |>
  summarize(trial_count = n(), .by=c(has_cdi, subject_id, dataset_name))

dataset_subject_counts <- subject_counts |>
  summarize(subject_count = n(), average_trial_count = mean(trial_count), cdi_count = sum(has_cdi == TRUE), non_cdi_count = (sum(has_cdi == FALSE)), .by=dataset_name)  |>
  pivot_longer(cols = c(cdi_count, non_cdi_count), 
               names_to = "cdi", 
               values_to = "count")

# Create stacked bar chart
ggplot(dataset_subject_counts, aes(x = reorder(paste0(dataset_name,"(subj trials M=", round(average_trial_count, 1), ")"), -count), y = count)) +
  geom_bar(stat = "identity", position = "stack", fill="skyblue") +
  labs(title = "Number of participants",
       x = "Dataset Name", 
       y = "Count") +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

#### Per trial
```{r}
trial_participant_counts <- usable_trials_summarized_with_sims |> summarize(
  admin_count = n(), .by=c(trial_type_id, dataset_name)
)

admin_based_trials <- trial_participant_counts |> mutate(
  rounded_count = round_to_nearest(admin_count, round_to=5),
  rounded_count = ifelse(rounded_count > 100, 100, rounded_count)
)

clip_admin_stacked_bar <- admin_based_trials |>
  #filter(N > 10) |>
  group_by(rounded_count, dataset_name) |>
  summarize(trial_count = n(), .groups = "drop")

ggplot(clip_admin_stacked_bar, aes(x = factor(rounded_count), y = trial_count, fill = dataset_name)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Number of Trials by dataset and admins",
    x = "Number of participants by Trial type and Dataset (rounded to nearest 5 and clipped at 100)",
    y = "Number of trials",
    fill = "Dataset"
  ) +
  theme_minimal()
```

