---
title: "Peekbank similarity analyses"
author: "Tarun Sepuri"
date: "`r format(Sys.time(), '%d %B, %Y')`"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    self-contained: true
execute:
  warning: false
  message: false
---

```{r}
library(tidyverse)
source("helpers.R")
library(ggrepel)
library(cowplot)
library(here)
```

# Import data 
```{r}
usable_trials_summarized_with_sims <- read.csv(here("data/usable_trials_with_similarities.csv"))
aoa_ratings <- read.csv(here("data/metadata/level-aoaratings_type-kuperman_data.csv"))
saliency_values <- read.csv(here("data/metadata/level-imagepair_added-saliency_data.csv"))
usable_trials_summarized_with_sims <- usable_trials_summarized_with_sims |> 
  left_join(aoa_ratings |> transmute(text1=Word, aoa=AoA_Kup_lem)) |>
  left_join(saliency_values, by=c("unique_pair"="ImagePair")) |>
  mutate(age_bucket = cut(age, 
                          breaks = c(12, 24, 36, 48, 60), 
                          labels = c("(12-24]", "(24-36]", "(36-48]", "(48-60]"),
                          include.lowest = TRUE))

# rounding each participant to the closest 5
age_based_trials <- usable_trials_summarized_with_sims |> mutate(
  rounded_age = round_to_nearest(age, round_to=5)
)

clip_data_summarized <- summarize_similarity_data_collapsed(usable_trials_summarized_with_sims, extra_fields = c("dataset_name", "vanilla_trial", "aoa", "MeanSaliencyDiff")) |> mutate(
   sim_logit = qlogis(pmin(pmax(image_similarity, 1e-6), 1 - 1e-6))
)

clip_data_summarized_age_bucket <- summarize_similarity_data_collapsed(usable_trials_summarized_with_sims, extra_fields = c("dataset_name", "vanilla_trial", "aoa", "MeanSaliencyDiff", "age_bucket")) |> mutate(
   sim_logit = qlogis(pmin(pmax(image_similarity, 1e-6), 1 - 1e-6))
)
```

```{r}
 clip_data_summarized |> mutate(
    sim_bucket = cut(
      image_similarity,
      breaks = seq(0, 1, by = 0.05),
      include.lowest = TRUE,
      right = FALSE
    )
  ) |>
  count(sim_bucket) |>
  ggplot(aes(x = n, y = sim_bucket)) +
  geom_col() +
  labs(
    x = "Count",
    y = "Image similarity bucket",
    title = "Distribution of image similarity"
  ) +
  theme_minimal()
```

# Main similarity plots

CLIP analysis: current similarity effects are dubious with lots of dataset level variance of course that will have to be accounted for in a mixed effects model; Garrison Bergelson dataset has individualized trials so difficult to make use of as well.

```{r fig.width=15, fig.height=15}
# N = the number of participants in a single trial here
adams_marchman_data_summarized <- summarize_similarity_data_collapsed(usable_trials_summarized_with_sims, extra_fields = c("dataset_name", "vanilla_trial")) |> filter(N > 10 & dataset_name == "adams_marchman_2018" & vanilla_trial==1)
am_plots <- generate_multimodal_plots(adams_marchman_data_summarized, "CLIP", title="Adams & Marchman, 2018")
am_plots

weaver_zettersten_data_summarized <- summarize_similarity_data_collapsed(usable_trials_summarized_with_sims, extra_fields = c("dataset_name", "vanilla_trial")) |> filter(N > 10 & dataset_name == "weaver_zettersten_2024" & vanilla_trial==1)
wz_plots <- generate_multimodal_plots(weaver_zettersten_data_summarized, "CLIP", title="Weaver et al., 2024")
wz_plots

clip_plots <- generate_multimodal_plots(clip_data_summarized |> filter(vanilla_trial==1), "CLIP", title="all vanilla trials")
clip_plots
```

## more multimodal plot visualizations
```{r fig.width=15, fig.height=15}
vanilla <- clip_data_summarized |> filter(vanilla_trial == 1)

generate_multimodal_plots(vanilla, "CLIP", title = "all vanilla trials", size_by_n = TRUE, base_size=16)
generate_multimodal_plots(vanilla, "CLIP", title = "by dataset",   group_var = "dataset_name", se_alpha=0.2, point_alpha=0.1, show_legend = TRUE, line_alpha=0.4, show_cor=FALSE, overall_line=TRUE, overall_alpha=0.7, base_size=16)
generate_multimodal_plots(clip_data_summarized, "CLIP", title = "by trial type", group_var = "vanilla_trial", point_alpha=0.1, show_se=TRUE, base_size=16)
generate_multimodal_plots(clip_data_summarized_age_bucket |> filter(vanilla_trial == 1 & !is.na(age_bucket)), "CLIP", title = "different age buckets", group_var = "age_bucket",  show_se = TRUE, se_alpha = 0.2, point_alpha=0.1, line_alpha = 0.8, line_width=1.1, base_size=16)
```

## comparing all IVs
```{r}
library(GGally)
confusability_vars <- clip_data_summarized |>
  filter(vanilla_trial==1) |>
  transmute(
    mean_value=scale(mean_value)[, 1],
    image_similarity=scale(image_similarity)[, 1],
    text_similarity=scale(text_similarity)[, 1],
    multimodal_similarity=scale(multimodal_similarity)[, 1],
    ooo_similarity=scale(ooo_similarity)[, 1],
    aoa=scale(aoa)[, 1],
    MeanSaliencyDiff=scale(MeanSaliencyDiff)[, 1]
  )

p <- ggpairs(
  confusability_vars,
  upper = list(continuous = wrap("cor", method = "spearman",
                                 use = "pairwise.complete.obs", size = 3)),
  lower = list(continuous = wrap("smooth", method = "lm",
                                 alpha = 0.4, size = 0.8)),
  diag  = list(continuous = wrap("densityDiag", alpha = 0.5))
) +
  theme_bw()

p
```
lots of interesting colinear effects here with AoA and mean saliency..


# Analysis across age
```{r}
calculate_correlations <- function(data, x_var, y_var, group_var = c("rounded_age"), conf_level = 0.95) {
  data |>
    group_by(across(all_of(group_var))) |>
    summarize(
      {
        cor_test <- cor.test(.data[[x_var]], .data[[y_var]], method = "pearson", conf.level = conf_level)
        tibble(
          pearson_cor = cor_test$estimate,
          p_value = cor_test$p.value,
          ci_lower = cor_test$conf.int[1],
          ci_upper = cor_test$conf.int[2]
        )
      },
      .groups = "drop"
    )
}

sim_age_plot <- function(data) {
  ggplot(data, aes(x = rounded_age, y = pearson_cor, color = similarity_type)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    geom_point(size = 3, position = position_dodge(width=0.5)) +  # Apply jitter to points only
    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), 
                  width = 0.3, alpha = 0.2,
                  position=position_dodge(width=0.5)) +  # No jitter on error bars
    geom_smooth(span = 2, alpha=0.1, se=FALSE) +
    labs(title = paste("Similarity correlations across age"),
         x = "Age",
         y = "Coefficient of similarity") +  
    theme_minimal() +
    guides(shape = "none") +
    scale_x_continuous(breaks=seq(5,70,5)) +
    scale_color_brewer(palette = "Set1", name = "Similarity type") 
}

# can't figure out what values to filter to here.
clip_data_age_summarized <- summarize_similarity_data_collapsed(age_based_trials |> filter(vanilla_trial == 1), extra_fields = c("rounded_age", "dataset_name")) |> filter(N >= 5) |> group_by(rounded_age) |>
  filter(n() >= 5) |>
  ungroup()
clip_age_image_cors <- calculate_correlations(clip_data_age_summarized, "image_similarity", "mean_value") |> mutate(similarity_type = "image")
clip_age_text_cors <- calculate_correlations(clip_data_age_summarized, "text_similarity", "mean_value") |> mutate(similarity_type = "text")
clip_age_multimodal_cors <- calculate_correlations(clip_data_age_summarized, "multimodal_similarity", "mean_value") |> mutate(similarity_type = "multimodal")
clip_age_ooo_cors <- calculate_correlations(clip_data_age_summarized, "ooo_similarity", "mean_value") |> mutate(similarity_type = "ooo")
clip_age_cors <- bind_rows(clip_age_image_cors, clip_age_text_cors, clip_age_multimodal_cors, clip_age_ooo_cors)
sim_age_plot(clip_age_cors)

ggplot(clip_age_cors, aes(x = rounded_age, y = pearson_cor)) +
  geom_point(aes(color = p_value < 0.05), size = 3) + 
  geom_smooth(span = 2) +
  labs(title = "Image similarity correlation across age",
       x = "Age in months",
       y = "Pearson Correlation") +
  scale_color_manual(values = c("TRUE" = "black", "FALSE" = "gray")) +  # Set color for significance
  theme_minimal() +
  theme(legend.position = "none")
```


# stats

## pre-registered models
```{r}
library(lmerTest)
library(glmmTMB)
library(MuMIn)
library(broom.mixed)
model_data <- usable_trials_summarized_with_sims |> filter(vanilla_trial == 1)
sims <- c("image_similarity","text_similarity","multimodal_similarity","ooo_similarity")
fit_main <- function(sim,
                     data = model_data,
                     response = "mean_target_looking_critical_window",
                     added_structure = NULL, pruned_model=FALSE) {
  terms <- c(
    sprintf("scale(%s)*scale(age)", sim),
    "(1 | dataset_id)"
  )
  if (pruned_model) {
    terms <- c(terms, "(1 | subject_id)")
  } else {
    terms <- c(terms, sprintf("(1 + scale(%s) | subject_id)", sim))
  }
  if (!is.null(added_structure)) {
    terms <- c(terms, added_structure)
  }
  f <- reformulate(
    terms,
    response = sprintf("scale(%s)", response)
  )
  lmer(f, data = data)
}

mods <- lapply(sims, fit_main); names(mods) <- sims
lapply(mods, function(m) m@optinfo$conv$lme4$messages)
lapply(mods, function(m) summary(m))
r.squaredGLMM(mods$image_similarity)
```
Trying to see why text and multimodal similarity models were singular.
```{r}
vars <- c("image_similarity", "text_similarity", "multimodal_similarity",
          "ooo_similarity", "mean_target_looking_critical_window",
          "age", "aoa", "MeanSaliencyDiff")

colSums(is.na(model_data[vars]))

# total rows, and rows complete across all vars 
nrow(model_data)
sum(complete.cases(model_data[vars]))

# rows complete without requiring ooo
sum(complete.cases(model_data[setdiff(vars, "ooo_similarity")]))
```
### getting rid of singular effects
getting rid of singular effects 
```{r}
pruned_text_model <- lmer(scale(mean_target_looking_critical_window) ~ scale(text_similarity)*scale(age)
                    + scale(aoa) + scale(MeanSaliencyDiff)
                    + (1 | subject_id)
                    + (1 | text1:unique_pair)
                    + (1 | dataset_id), data = model_data)

pruned_multimodal_model <- lmer(scale(mean_target_looking_critical_window) ~ scale(multimodal_similarity)*scale(age)
                    + scale(aoa) + scale(MeanSaliencyDiff)
                    + (1 | subject_id)
                    + (1 | text1:unique_pair)
                    + (1|dataset_id), data = model_data)

summary(pruned_text_model)
summary(pruned_multimodal_model)
```

### model comparison
Effects persist in the pruned models. Now trying to compare to see which one has the best fit.
```{r}
dat_cc <- model_data |>
  tidyr::drop_na(image_similarity, text_similarity, multimodal_similarity, ooo_similarity,
                 mean_target_looking_critical_window,
                 age, aoa, MeanSaliencyDiff)

fit_common <- function(sim, data=dat_cc) {
  f <- reformulate(
    c(sprintf("scale(%s)*scale(age)", sim),
      "scale(aoa)", "scale(MeanSaliencyDiff)",
      "(1 | administration_id)",
      "(1 | text1:unique_pair)",
      "(1 | dataset_id)"),
    response = "scale(mean_target_looking_critical_window)")
  lmer(f, data = dat_cc, REML = FALSE)
}

mods_common <- lapply(sims, fit_common); names(mods_common) <- sims

sel <- model.sel(mods_common$image_similarity, mods_common$text_similarity,
                 mods_common$multimodal_similarity, mods_common$ooo_similarity)
sel

summary(mods_common$image_similarity)
```
Image similarity model comes out on top.

## including non-vanilla trials
```{r}
mods_all <- lapply(sims, fit_main, usable_trials_summarized_with_sims); names(mods_all) <- sims
lapply(mods_all, function(m) m@optinfo$conv$lme4$messages)
lapply(mods_all, function(m) summary(m))
```

multimodal similarity and image similarity ar  the only significant ones here...pruning
```{r}
pruned_multimodal_model_all <- lmer(scale(mean_target_looking_critical_window) ~ scale(multimodal_similarity)*scale(age)
                    + scale(aoa) + scale(MeanSaliencyDiff)
                    + (1 | subject_id)
                    + (1|dataset_id), data = usable_trials_summarized_with_sims)
summary(pruned_multimodal_model_all)
```
effect stays and with a pretty robust random effects structure and with an interaction with age. feels like this is a really interesting effect, especially given that the other models across vanilla and non-vanilla are not similarly predictive.

## including baseline window as covariate
```{r}
baseline_data <- model_data |> filter(min_time <= -500)

mods_baseline_covariate <-  lapply(sims, fit_main, data=baseline_data, added_structure="mean_target_looking_baseline_window", pruned_model=TRUE); names(mods_baseline_covariate) <- sims
lapply(mods_baseline_covariate, function(m) summary(m))
```

## predicting baseline corrected looking
```{r}
mods_baseline_corrected <-  lapply(sims, fit_main, data=baseline_data, response="corrected_target_looking", pruned_model=TRUE); names(mods_baseline_corrected) <- sims
lapply(mods_baseline_corrected, function(m) summary(m))
```

## Under 3 years old
LWL specific ages and AoA?
```{r}
mods_younger <-  lapply(sims, fit_main, data=model_data |> filter(age < 36), response="corrected_target_looking", pruned_model=TRUE); names(mods_younger) <- sims
lapply(mods_younger, function(m) summary(m))
```
AoA effect is significant here which makes some intuitive sense.