---
title: "Peekbank similarity analyses"
author: "Tarun Sepuri"
date: "`r format(Sys.time(), '%d %B, %Y')`"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    self-contained: true
execute:
  warning: false
  message: false
---

```{r}
library(tidyverse)
source("helpers.R")
library(ggrepel)
library(cowplot)
library(here)
library(colorspace)
library(ggeffects)
```

# Import data 
```{r}
usable_trials_summarized_with_sims <- read.csv(here("data/usable_trials_with_similarities.csv"))
aoa_ratings <- read.csv(here("data/metadata/level-aoaratings_type-kuperman_data.csv"))
saliency_values <- read.csv(here("data/metadata/level-imagepair_added-saliency_data.csv"))
usable_trials_summarized_with_sims <- usable_trials_summarized_with_sims |> 
  left_join(aoa_ratings |> transmute(text1=Word, aoa=AoA_Kup_lem)) |>
  left_join(saliency_values, by=c("unique_pair"="ImagePair")) |>
  mutate(age_bucket = cut(age, 
                          breaks = c(12, 18, 24, 36, 48, 60), 
                          labels = c("(12-18]", "(18-24]", "(24-36]", "(36-48]", "(48-60]"),
                          include.lowest = TRUE))

# rounding each participant to the closest 5
age_based_trials <- usable_trials_summarized_with_sims |> mutate(
  rounded_age = round_to_nearest(age, round_to=5)
)

clip_data_summarized <- summarize_similarity_data_collapsed(usable_trials_summarized_with_sims, extra_fields = c("dataset_name", "vanilla_trial", "aoa", "MeanSaliencyDiff")) |> mutate(
   sim_logit = qlogis(pmin(pmax(image_similarity, 1e-6), 1 - 1e-6))
)

clip_data_summarized_age_bucket <- summarize_similarity_data_collapsed(usable_trials_summarized_with_sims, extra_fields = c("dataset_name", "vanilla_trial", "aoa", "MeanSaliencyDiff", "age_bucket")) |> mutate(
   sim_logit = qlogis(pmin(pmax(image_similarity, 1e-6), 1 - 1e-6))
)

vanilla <- clip_data_summarized |> filter(vanilla_trial == 1)

sims <- c("image_similarity","text_similarity","multimodal_similarity","ooo_similarity")
facet_labs <- c(text_similarity       = "Text similarity",
                image_similarity      = "Image similarity",
                multimodal_similarity = "Multimodal similarity",
                ooo_similarity        = "Adult behavioral similarity")

downloaded_aois <- readRDS(here("intermediates/downloaded_aois.Rds"))
```

```{r}
 clip_data_summarized |> mutate(
    sim_bucket = cut(
      image_similarity,
      breaks = seq(0, 1, by = 0.05),
      include.lowest = TRUE,
      right = FALSE
    )
  ) |>
  count(sim_bucket) |>
  ggplot(aes(x = n, y = sim_bucket)) +
  geom_col() +
  labs(
    x = "Count",
    y = "Image similarity bucket",
    title = "Distribution of image similarity"
  ) +
  theme_minimal()
```

# Main similarity plots

CLIP analysis: current similarity effects are dubious with lots of dataset level variance of course that will have to be accounted for in a mixed effects model; Garrison Bergelson dataset has individualized trials so difficult to make use of as well.

```{r fig.width=15, fig.height=15}
# N = the number of participants in a single trial here
adams_marchman_data_summarized <- summarize_similarity_data_collapsed(usable_trials_summarized_with_sims, extra_fields = c("dataset_name", "vanilla_trial")) |> filter(N > 10 & dataset_name == "adams_marchman_2018" & vanilla_trial==1)
am_plots <- generate_multimodal_plots(adams_marchman_data_summarized, "CLIP", title="Adams & Marchman, 2018")
am_plots

weaver_zettersten_data_summarized <- summarize_similarity_data_collapsed(usable_trials_summarized_with_sims, extra_fields = c("dataset_name", "vanilla_trial")) |> filter(N > 10 & dataset_name == "weaver_zettersten_2024" & vanilla_trial==1)
wz_plots <- generate_multimodal_plots(weaver_zettersten_data_summarized, "CLIP", title="Weaver et al., 2024")
wz_plots

clip_plots <- generate_multimodal_plots(clip_data_summarized |> filter(vanilla_trial==1), "CLIP", title="all vanilla trials")
clip_plots
```

```{r fig.width=15, fig.height=15}
generate_multimodal_plots(vanilla |> filter(N >= 5), "CLIP", title = "all vanilla trials (N >= 5)", size_by_n = TRUE, base_size=16)
```

## more multimodal plot visualizations
```{r fig.width=15, fig.height=15}

generate_multimodal_plots(vanilla, "CLIP", title = "all vanilla trials", size_by_n = TRUE, base_size=16)
generate_multimodal_plots(vanilla, "CLIP", title = "by dataset",   group_var = "dataset_name", se_alpha=0.2, point_alpha=0.1, show_legend = TRUE, line_alpha=0.4, show_cor=FALSE, overall_line=TRUE, overall_alpha=0.7, base_size=16)
generate_multimodal_plots(clip_data_summarized, "CLIP", title = "different trial types", group_var = "vanilla_trial", point_alpha=0.1, show_se=TRUE, base_size=16)
generate_multimodal_plots(clip_data_summarized_age_bucket |> filter(vanilla_trial == 1 & !is.na(age_bucket) & N>=5), "CLIP", title = "different age buckets (N>=5)", group_var = "age_bucket",  show_se = TRUE, se_alpha = 0.2, point_alpha=0.1, line_alpha = 0.8, line_width=1.1, base_size=16)
```

## By age
```{r}
plot_data_age <- clip_data_summarized_age_bucket |>
  pivot_longer(all_of(sims), names_to = "sim_type", values_to = "similarity") |>
  mutate(sim_type = factor(sim_type, levels = sims)) 

ggplot(
  plot_data_age |> filter(!is.na(age_bucket)),
  aes(
    similarity,
    mean_value,
    colour = age_bucket,
    fill = age_bucket
  )
) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  geom_point(alpha = 0.1, size = 2) +
  
  stat_smooth(
    method = "lm",
    se = TRUE,
    aes(fill = age_bucket),
    alpha = 0.15   # ribbon alpha
  ) +

  stat_smooth(
    method = "lm",
    se = FALSE,
    linewidth = 1.2
  ) +
  facet_wrap(
    ~ sim_type,
    nrow = 2,
    scales = "free_x",
    labeller = labeller(sim_type = facet_labs),
    strip.position = "top"
  ) +
  scale_color_discrete_qualitative(palette = "Set 3") +
  labs(
    x = "Target–distractor similarity",
    y = "Proportion target looking in critical window",
    colour = "Age bucket"
  ) +
  guides(fill = "none") +
  theme_classic(base_size = 16) +
  theme(
    legend.position = "bottom",
    strip.text = element_text(size = 20),
    panel.spacing = unit(1, "lines"),
    plot.margin = margin(t = 5.5, r = 30, b = 5.5, l = 5.5, unit = "pt")
  )
```

## comparing all IVs
```{r}
library(GGally)
confusability_vars <- clip_data_summarized |>
  filter(vanilla_trial==1) |>
  transmute(
    mean_value=scale(mean_value)[, 1],
    image_similarity=scale(image_similarity)[, 1],
    text_similarity=scale(text_similarity)[, 1],
    multimodal_similarity=scale(multimodal_similarity)[, 1],
    ooo_similarity=scale(ooo_similarity)[, 1],
    aoa=scale(aoa)[, 1],
    MeanSaliencyDiff=scale(MeanSaliencyDiff)[, 1]
  )

p <- ggpairs(
  confusability_vars,
  upper = list(continuous = wrap("cor", method = "spearman",
                                 use = "pairwise.complete.obs", size = 3)),
  lower = list(continuous = wrap("smooth", method = "lm",
                                 alpha = 0.4, size = 0.8)),
  diag  = list(continuous = wrap("densityDiag", alpha = 0.5))
) +
  theme_bw()

p
```
so many interesting colinear effects here with AoA and mean saliency..

```{r}
# Load required libraries
library(ggcorrplot)
library(Hmisc) # Required for computing p-values

# 1. Prepare and rename your data
confusability_vars <- clip_data_summarized |>
  filter(vanilla_trial == 1) |>
  transmute(
  #  mean_value = scale(mean_value)[, 1],
    image_similarity = scale(image_similarity)[, 1],
    text_similarity = scale(text_similarity)[, 1],
    multimodal_similarity = scale(multimodal_similarity)[, 1],
    ooo_similarity = scale(ooo_similarity)[, 1],
    aoa = scale(aoa)[, 1],
    MeanSaliencyDiff = scale(MeanSaliencyDiff)[, 1]
  ) |>
  # RENAME YOUR PREDICTORS HERE:
  # Format: "New Name" = old_name
  rename(
   # "Critical window looking"            = mean_value,
    "Image similarity"      = image_similarity,
    "Text similarity"       = text_similarity,
    "Multimodal similarity" = multimodal_similarity,
    "Adult behavioral similarity"        = ooo_similarity,
    "Estimated target age of acquisition"    = aoa,
    "Saliency difference"    = MeanSaliencyDiff
  )

cor_data <- rcorr(as.matrix(confusability_vars), type = "spearman")

# 3. Flatten matrices into a long dataframe for ggplot
r_df <- as.data.frame(cor_data$r) %>% 
  rownames_to_column(var = "Var1") %>% 
  pivot_longer(-Var1, names_to = "Var2", values_to = "r")

p_df <- as.data.frame(cor_data$P) %>% 
  rownames_to_column(var = "Var1") %>% 
  pivot_longer(-Var1, names_to = "Var2", values_to = "p")

# Combine r and p data, and create the significance star labels
plot_data <- left_join(r_df, p_df, by = c("Var1", "Var2")) %>%
  # Fill diagonal p-values (which are NA) with 1 so they don't trigger stars
  mutate(p = ifelse(Var1 == Var2, 1, p)) %>% 
  mutate(
    stars = case_when(
      p < 0.001 ~ "***",
      p < 0.01  ~ "**",
      p < 0.05  ~ "*",
      TRUE      ~ ""
    ),
    # Combine the correlation number and stars into one text label
    label_text = paste0(round(r, 2), stars)
  )

# 4. Filter to get only the LOWER half of the square matrix
# We convert variables to factors to respect the order of our dataframe
vars_order <- colnames(confusability_vars)
plot_data <- plot_data %>%
  mutate(
    Var1 = factor(Var1, levels = vars_order),
    Var2 = factor(Var2, levels = vars_order)
  ) %>%
  filter(as.integer(Var1) > as.integer(Var2)) # Removes diagonal and upper triangle

# 5. Build the Heatmap using ggplot2
ggplot(plot_data, aes(x = Var2, y = Var1, fill = r)) +
  geom_tile(color = "white") +
  # Add the correlation values and significance stars combined
  geom_text(aes(label = label_text), size = 3.5, color = "black") +
  # Define the diverging color palette (Negative, Neutral, Positive)
  scale_fill_gradient2(
    low = "#6D9EC1", mid = "white", high = "#E46726", 
    midpoint = 0, limit = c(-1, 1), space = "Lab", 
    name = "Spearman\ncorrelation"
  ) +
  theme_bw() +
  labs(title = "Correlation of looking behavior predictors", x = NULL, y = NULL) +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank()
  )
ggsave(here("figures/correlation.png"))
```

# Analysis across age
```{r}
library(dplyr)
calculate_correlations <- function(data, x_var, y_var, group_var = c("rounded_age"), conf_level = 0.95) {
  data |>
    dplyr::group_by(across(all_of(group_var))) |>
    dplyr::summarize(
      {
        cor_test <- cor.test(.data[[x_var]], .data[[y_var]], method = "pearson", conf.level = conf_level)
        tibble(
          pearson_cor = cor_test$estimate,
          p_value = cor_test$p.value,
          ci_lower = cor_test$conf.int[1],
          ci_upper = cor_test$conf.int[2]
        )
      },
      .groups = "drop"
    )
}

sim_age_plot <- function(data) {
  ggplot(data, aes(x = rounded_age, y = pearson_cor, color = similarity_type)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    geom_point(size = 3, position = position_dodge(width=0.5)) +  # Apply jitter to points only
    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), 
                  width = 0.3, alpha = 0.2,
                  position=position_dodge(width=0.5)) +  # No jitter on error bars
    geom_smooth(span = 2, alpha=0.1, se=FALSE) +
    labs(title = paste("Similarity correlations across age"),
         x = "Age",
         y = "Coefficient of similarity") +  
    theme_minimal() +
    guides(shape = "none") +
    scale_x_continuous(breaks=seq(5,70,5)) +
    scale_color_brewer(palette = "Set1", name = "Similarity type") 
}

# can't figure out what values to filter to here?
clip_data_age_summarized <- summarize_similarity_data_collapsed(age_based_trials |> filter(vanilla_trial == 1), extra_fields = c("rounded_age", "dataset_name")) |> filter(N >= 5) |> group_by(rounded_age) |>
  filter(n() >= 5) |>
  ungroup()
clip_age_image_cors <- calculate_correlations(clip_data_age_summarized, "image_similarity", "mean_value") |> mutate(similarity_type = "image")
clip_age_text_cors <- calculate_correlations(clip_data_age_summarized, "text_similarity", "mean_value") |> mutate(similarity_type = "text")
clip_age_multimodal_cors <- calculate_correlations(clip_data_age_summarized, "multimodal_similarity", "mean_value") |> mutate(similarity_type = "multimodal")
clip_age_ooo_cors <- calculate_correlations(clip_data_age_summarized, "ooo_similarity", "mean_value") |> mutate(similarity_type = "ooo")
clip_age_cors <- bind_rows(clip_age_image_cors, clip_age_text_cors, clip_age_multimodal_cors, clip_age_ooo_cors)
sim_age_plot(clip_age_cors)

ggplot(clip_age_cors, aes(x = rounded_age, y = pearson_cor)) +
  geom_point(aes(color = p_value < 0.05), size = 3) + 
  geom_smooth(span = 2) +
  labs(title = "Image similarity correlation across age",
       x = "Age in months",
       y = "Pearson Correlation") +
  scale_color_manual(values = c("TRUE" = "black", "FALSE" = "gray")) +  # Set color for significance
  theme_minimal() +
  theme(legend.position = "none")
```
# time
```{r}
library(tidyr)
crit <- qnorm(0.975)

# one row per trial type, median-split within each similarity dimension
sims_long <- usable_trials_summarized_with_sims |> filter(vanilla_trial == 1) |>
  select(trial_type_id, ends_with("_similarity")) |>
  distinct() |>
  pivot_longer(ends_with("_similarity"),
               names_to = "sim_type", values_to = "similarity") |>
  filter(!is.na(similarity)) |>
  group_by(sim_type) |>
  mutate(sim_split = if_else(similarity >= median(similarity, na.rm = TRUE),
                             "high", "low")) |>
  ungroup() |>
  mutate(sim_type = sub("_similarity$", "", sim_type))

aoi_base <- downloaded_aois |>
  filter(!is.na(accuracy), trial_type_id %in% sims_long$trial_type_id)

# fan out: each timepoint contributes to its high/low bucket in each dimension
aoi_split <- aoi_base |>
  left_join(sims_long, by = "trial_type_id", relationship = "many-to-many")

# subject means per condition x time, then across-subject mean + 95% CI
across_split <- aoi_split |>
  group_by(sim_type, sim_split, administration_id, t_norm) |>
  summarise(mean_accuracy = mean(accuracy, na.rm = TRUE), .groups = "drop") |>
  group_by(sim_type, sim_split, t_norm) |>
  summarise(accuracy = mean(mean_accuracy, na.rm = TRUE),
            ci = crit * sd(mean_accuracy, na.rm = TRUE) / sqrt(n()),
            .groups = "drop") |>
  mutate(condition = paste(sim_type, sim_split))

summarize_across_subj_aois <- readRDS(here("intermediates/summarize_across_subj_aois.Rds"))
# the existing overall line as the 9th condition
plot_df <- bind_rows(
  across_split |> select(t_norm, accuracy, ci, condition),
  summarize_across_subj_aois |> mutate(condition = "Overall") |>
    select(t_norm, accuracy, ci, condition)
) |>
  mutate(condition = forcats::fct_relevel(condition, "Overall"))

looking_times <- ggplot(plot_df, aes(t_norm, accuracy,
                                      colour = condition, fill = condition)) +
  xlim(-2000, 4000) +
  geom_errorbar(aes(ymin = accuracy - ci, ymax = accuracy + ci),
                width = 0, alpha = 0.15) +
  geom_smooth(method = "gam", se = FALSE, linewidth = 1) +
  geom_vline(xintercept = 0, size = 1.5) +
  geom_hline(yintercept = 0.5, size = 1.2, linetype = "dashed") +
  geom_vline(xintercept = 300, linetype = "dotted") +
  ylim(0, 1) +
  xlab("Time (normalized to target word onset) in ms") +
  ylab("Proportion Target Looking") +
  labs(colour = "Condition", fill = "Condition")
looking_times
```

# stats

## pre-registered models
```{r}
library(lmerTest)
library(glmmTMB)
library(MuMIn)
library(broom.mixed)
model_data <- usable_trials_summarized_with_sims |> filter(vanilla_trial == 1)
model_data_scaled <- model_data |> 
  mutate(
    # Scale your response variable but keep the original for plotting later
    scaled_response = as.vector(scale(mean_target_looking_critical_window)),
    scaled_age = as.vector(scale(age)),
    scaled_aoa = as.vector(scale(aoa)),
    scaled_saliency_diff = as.vector(scale(MeanSaliencyDiff))
  ) |> 
  # Dynamic scaling for your various similarity columns
  mutate(across(all_of(sims), ~as.vector(scale(.x)), .names = "scaled_{.col}"))

fit_main <- function(sim,
                     data = model_data,
                     response = "mean_target_looking_critical_window",
                     added_structure = NULL, pruned_model=FALSE) {
  terms <- c(
    sprintf("scale(%s)*scale(age)", sim),
    "scale(aoa)",
    "scale(MeanSaliencyDiff)",
    "(1 | original_target_label:img_key)",
    "(1 | dataset_id)"
  )
  if (pruned_model) {
    terms <- c(terms, "(1 | subject_id)")
  } else {
    terms <- c(terms, sprintf("(1 + scale(%s) | subject_id)", sim))
  }
  if (!is.null(added_structure)) {
    terms <- c(terms, added_structure)
  }
  f <- reformulate(
    terms,
    response = sprintf("scale(%s)", response)
  )
  lmer(f, data = data)
}

fit_main_scaled <- function(sim, data = model_data_scaled, pruned_model = FALSE) {
  
  # Use the pre-scaled variable names
  terms <- c(
    sprintf("scaled_%s * scaled_age", sim),
    "scaled_aoa",
    "scaled_saliency_diff",
    "(1 | original_target_label:img_key)",
    "(1 | dataset_id)"
  )
  
  if (pruned_model) {
    terms <- c(terms, "(1 | subject_id)")
  } else {
    terms <- c(terms, sprintf("(1 + scaled_%s | subject_id)", sim))
  }
  
  # Response is already pre-scaled
  f <- reformulate(terms, response = "scaled_response")
  
  lmer(f, data = data)
}

mods <- lapply(sims, fit_main); names(mods) <- sims
mods_scaled <- lapply(sims, fit_main_scaled); names(mods_scaled) <- sims
lapply(mods, function(m) m@optinfo$conv$lme4$messages)
lapply(mods, function(m) summary(m))
r.squaredGLMM(mods$image_similarity)
```

Trying to see why text and multimodal similarity models were singular.
```{r}
vars <- c("image_similarity", "text_similarity", "multimodal_similarity",
          "ooo_similarity", "mean_target_looking_critical_window",
          "age", "aoa", "MeanSaliencyDiff")

colSums(is.na(model_data[vars]))

# total rows, and rows complete across all vars 
nrow(model_data)
sum(complete.cases(model_data[vars]))

# rows complete without requiring ooo
sum(complete.cases(model_data[setdiff(vars, "ooo_similarity")]))
```

```{r}
library(ggeffects)
mu  <- mean(model_data$mean_target_looking_critical_window, na.rm = TRUE)
sig <- sd(model_data$mean_target_looking_critical_window, na.rm = TRUE)

all_predictions <- lapply(sims, function(sim) {
  term_name <- sim
  
  # Generate prediction
  pred <- ggpredict(mods[[sim]], terms = term_name)
  
  # Label it so we know which facet it belongs to
  pred$sim_type <- paste(gsub("_", " ", sim))
  return(pred)
})

combined_main_preds <- do.call(rbind, all_predictions)

combined_main_preds$predicted_raw <- combined_main_preds$predicted * sig + mu
combined_main_preds$conf.low_raw  <- combined_main_preds$conf.low  * sig + mu
combined_main_preds$conf.high_raw <- combined_main_preds$conf.high * sig + mu

ggplot(
  combined_main_preds, 
  aes(x = x, y = predicted_raw)
) +
  # Confidence intervals ribbon
  geom_ribbon(
    aes(ymin = conf.low_raw, ymax = conf.high_raw), 
    alpha = 0.15, 
    fill = "steelblue"
  ) +
  # Main effect prediction line
  geom_line(
    linewidth = 1.2, 
    color = "steelblue"
  ) +
  # 50% chance line (useful for target looking paradigms)
  geom_hline(
    yintercept = 0.5, 
    linetype = "dashed", 
    alpha = 0.5
  ) +
  # Separate into 4 panels side-by-side / wrapped
  facet_wrap(~ sim_type, scales = "free_x") +
  labs(
    title = "Main similarity predictors",
    subtitle = "Critical Window Predictions",
    x = "Target-distractor similarity",
    y = "Predicted Proportion Target Looking"
  ) +
  theme_classic() +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 11)
  )
```

### getting rid of singular effects
getting rid of singular effects 
```{r}
pruned_text_model <- lmer(scale(mean_target_looking_critical_window) ~ scale(text_similarity)*scale(age)
                    + scale(aoa) + scale(MeanSaliencyDiff)
                    + (1 | subject_id)
                    + (1 | original_target_label:img_key)
                    + (1 | dataset_id), data = model_data)

summary(pruned_text_model)
```

### model comparison
Effects persist in the pruned models. Now trying to compare to see which one has the best fit.
```{r}
dat_cc <- model_data |>
  tidyr::drop_na(image_similarity, text_similarity, multimodal_similarity, ooo_similarity,
                 mean_target_looking_critical_window,
                 age, aoa, MeanSaliencyDiff)

fit_common <- function(sim, data=dat_cc) {
  f <- reformulate(
    c(sprintf("scale(%s)*scale(age)", sim),
      "scale(aoa)", "scale(MeanSaliencyDiff)",
      "(1 | subject_id)",
      "(1 | original_target_label:img_key)",
      "(1 | dataset_id)"),
    response = "scale(mean_target_looking_critical_window)")
  lmer(f, data = dat_cc, REML = FALSE)
}

mods_common <- lapply(sims, fit_common); names(mods_common) <- sims

sel <- model.sel(mods_common$image_similarity, mods_common$text_similarity,
                 mods_common$multimodal_similarity, mods_common$ooo_similarity)
sel

summary(mods_common$image_similarity)
```
Image similarity model comes out on top.

## including non-vanilla trials
```{r}
model_data_all <- usable_trials_summarized_with_sims |>
  mutate(vanilla_trial = ifelse(vanilla_trial==1, 0.5, -0.5))
mods_all <- lapply(sims, fit_main, model_data_all, added_structure="scale(vanilla_trial)"); names(mods_all) <- sims
lapply(mods_all, function(m) m@optinfo$conv$lme4$messages)
lapply(mods_all, function(m) summary(m))
```

multimodal similarity and image similarity ar  the only significant ones here...pruning
```{r}
pruned_multimodal_model_all <- lmer(scale(mean_target_looking_critical_window) ~ scale(multimodal_similarity)*scale(age)
                    + scale(aoa) + scale(MeanSaliencyDiff)
                    + (1 | subject_id)
                    + (1|dataset_id), data = usable_trials_summarized_with_sims)
summary(pruned_multimodal_model_all)
```
effect stays and with a pretty robust random effects structure and with an interaction with age. feels like this is a really interesting effect, especially given that the other models across vanilla and non-vanilla are not similarly predictive.

## including baseline window as covariate
```{r}
baseline_data <- model_data |> filter(min_time <= -500) # getting rid of 5 individual trials ha

mods_baseline_covariate <-  lapply(sims, fit_main, data=baseline_data, added_structure="scale(mean_target_looking_baseline_window)", pruned_model=TRUE); names(mods_baseline_covariate) <- sims
lapply(mods_baseline_covariate, function(m) summary(m))
```

## predicting baseline corrected looking
```{r}
mods_baseline_corrected <-  lapply(sims, fit_main, data=baseline_data, response="corrected_target_looking"); names(mods_baseline_corrected) <- sims
lapply(mods_baseline_corrected, function(m) m@optinfo$conv$lme4$messages)
lapply(mods_baseline_corrected, function(m) summary(m))
```
## predicting window type
```{r}
baseline_data_window_type <- baseline_data |> pivot_longer(cols=c(mean_target_looking_critical_window, mean_target_looking_baseline_window), names_to="window_type", values_to="target_looking") |>
  mutate(window_type = str_replace(window_type, "mean_target_looking_", "")) |>
  mutate(trial_window_c = case_when(
    window_type=="critical_window" ~ 0.5,
    window_type=="baseline_window" ~ -0.5))
mods_baseline_window_type <- lapply(sims, function(sim) {

  fit_main(
    sim = sim,
    data = baseline_data_window_type,
    response = "target_looking",
    pruned_model = TRUE,
    added_structure = sprintf("scale(trial_window_c) * scale(%s)", sim)
  )

}); 
names(mods_baseline_window_type) <- sims
lapply(mods_baseline_window_type, function(m) m@optinfo$conv$lme4$messages)
lapply(mods_baseline_window_type, function(m) summary(m))
```
I mean these are all interesting, sure.

```{r}
# 1. Define models and scale constants
mm    <- mods_baseline_window_type[[3]] 
ooo   <- mods_baseline_window_type[[4]] 
text  <- mods_baseline_window_type[[2]] 
image <- mods_baseline_window_type[[1]] 

mu  <- mean(baseline_data_window_type$target_looking, na.rm = TRUE)
sig <- sd(baseline_data_window_type$target_looking, na.rm = TRUE)

# 2. Generate predictions (using unique variable names)
pred_mm    <- ggpredict(mm, terms = c("multimodal_similarity", "trial_window_c [-0.5, 0.5]"))
pred_ooo   <- ggpredict(ooo, terms = c("ooo_similarity", "trial_window_c [-0.5, 0.5]"))
pred_text  <- ggpredict(text, terms = c("text_similarity", "trial_window_c [-0.5, 0.5]"))
pred_image <- ggpredict(image, terms = c("image_similarity", "trial_window_c [-0.5, 0.5]"))

# 3. Add a source/type label to each dataframe
pred_mm$sim_type    <- "Multimodal similarity"
pred_ooo$sim_type   <- "Adult behavioral similarity"
pred_text$sim_type  <- "Text similarity"
pred_image$sim_type <- "Image similarity"

# 4. Combine all predictions into one dataframe
combined_preds <- rbind(pred_mm, pred_ooo, pred_text, pred_image)

combined_preds$sim_type <- factor(
  combined_preds$sim_type,
  levels = c(
    "Image similarity",
    "Text similarity",
    "Multimodal similarity",
    "Adult behavioral similarity"
  )
)

# 5. Un-scale / Back-transform the predicted values and confidence intervals
combined_preds$predicted_raw <- combined_preds$predicted * sig + mu
combined_preds$conf.low_raw  <- combined_preds$conf.low  * sig + mu
combined_preds$conf.high_raw <- combined_preds$conf.high * sig + mu

# 6. Create the Faceted Plot
baseline_interaction_plot <- ggplot(
  combined_preds,
  aes(x = x, y = predicted_raw, colour = group, fill = group)
) +
  geom_ribbon(
    aes(ymin = conf.low_raw, ymax = conf.high_raw),
    alpha = 0.15,
    colour = NA
  ) +
  geom_line(linewidth = 1.2) +
  geom_hline(
    yintercept = 0.5,
    linetype = "dashed",
    alpha = 0.7
  ) +
  # This splits your plot into 4 facets based on similarity type
  facet_wrap(~ sim_type, scales = "free_x") + 
  labs(
    title = "Effect of similarities by trial window type",
    x = "Target-distractor similarity",
    y = "Predicted proportion target looking",
    colour = "Trial window",
    fill = "Trial window"
  ) +
  scale_color_discrete(
    labels = c("Baseline window", "Critical window")
  ) +
  scale_fill_discrete(
    labels = c("Baseline window", "Critical window")
  ) +
  theme_classic() +
  theme(
    strip.background = element_blank(), # Cleans up facet header backgrounds
    strip.text = element_text(face = "bold", size = 11), # Emphasizes facet titles
    legend.position = "bottom" # Optional: Moves legend to bottom for wider facet viewing
  )
ggsave(here("figures/baseline_interaction.png"), baseline_interaction_plot)
```

## baseline window looking?
```{r}
mods_baseline_looking <-  lapply(sims, fit_main, data=baseline_data, response="mean_target_looking_baseline_window"); names(mods_baseline_looking) <- sims
lapply(mods_baseline_looking, function(m) m@optinfo$conv$lme4$messages)
lapply(mods_baseline_looking, function(m) summary(m))
```

## Under 3 years old
LWL specific ages and AoA?
```{r}
mods_younger <-  lapply(sims, fit_main, data=model_data |> filter(age <= 24)); names(mods_younger) <- sims
lapply(mods_younger, function(m) m@optinfo$conv$lme4$messages)
lapply(mods_younger, function(m) summary(m))
```
this makes me think of the cascading age prediction..younger kids: image similarity, older kids: multimodal, even older kids: text only?

Adult behavioral similarity is significant here..AoA effects only appear between 24 and 36 months.


# Conference plots
todo: make sure models used are not singular

## unscaled
```{r}
plot_data <- vanilla |>
  pivot_longer(all_of(sims), names_to = "sim_type", values_to = "similarity") |>
  mutate(sim_type = factor(sim_type, levels = sims))   # controls panel order

# ---- lmer prediction lines, one stacked data frame ---------------------------
pred <- bind_rows(lapply(sims, function(s) {
  m  <- mods[[s]]
  mf <- model.frame(m)                      # only the rows lmer kept
  # raw similarity range as the model saw it (mf stores the scaled column)
  scol <- mf[[grep(paste0("scale\\(", s), names(mf))]]
  raw  <- as.vector(scol) * attr(scol, "scaled:scale") + attr(scol, "scaled:center")
  x    <- seq(min(raw), max(raw), length.out = 150)
  d  <- model_data
  nd <- data.frame(x,
                   age = mean(d$age, na.rm = TRUE),
                   aoa = mean(d$aoa, na.rm = TRUE),
                   MeanSaliencyDiff = mean(d$MeanSaliencyDiff, na.rm = TRUE))
  names(nd)[1] <- s

  pr <- predict(m, newdata = nd, re.form = NA, se.fit = TRUE)

  rcol <- mf[[1]]
  rc <- attr(rcol, "scaled:center"); rs <- attr(rcol, "scaled:scale")

  data.frame(sim_type = s, similarity = x,
             y     =  pr$fit                     * rs + rc, # un-zscoring
             lower = (pr$fit - 1.96 * pr$se.fit) * rs + rc,
             upper = (pr$fit + 1.96 * pr$se.fit) * rs + rc)
})) |> mutate(sim_type = factor(sim_type, levels = sims))

library(colorspace)

fig <- ggplot(plot_data, aes(similarity, mean_value,
                             colour = dataset_name, fill = dataset_name)) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  geom_point(size = 3, alpha = 0.1) +
  stat_smooth(method = "glm", geom = "line", alpha = 0.4, linewidth = 1) +
  # lmer population line + CI
  geom_ribbon(data = pred, inherit.aes = FALSE,
              aes(x = similarity, ymin = lower, ymax = upper),
              fill = "black", alpha = 0.15) +
  geom_line(data = pred, inherit.aes = FALSE,
            aes(x = similarity, y = y),
            colour = "black", alpha = 0.7, linewidth = 1.2) +
  facet_wrap(~ sim_type, nrow = 2, scales = "free_x",
             labeller = labeller(sim_type = facet_labs),
             strip.position = "top") +          # strips read as x-axis titles
  scale_y_continuous(breaks = seq(0, 1, 0.2)) +
  coord_cartesian(ylim = c(0, 1)) +
  scale_color_discrete_qualitative(palette="Set 2") +
  #scale_color_manual(values = colorRampPalette(brewer.pal(8, "Dark2"))(24)) +
  labs(x = "Target–distractor similarity",
       y = "Proportion target looking in critical window",
       colour = "Dataset Name", fill = "Dataset Name") +
  theme_classic(base_size=16) +
  theme(legend.position = "bottom",
        #strip.placement = "outside",
        strip.text = element_text(size = 20),
         panel.spacing = unit(1, "lines"),
        plot.margin = margin(t = 5.5, r = 30, b = 5.5, l = 5.5, unit = "pt"))

ggsave(here("figures", "dataset_similarities.png"),
       fig, width = 16, height = 13, bg = "white")
fig
```

```{r}
library(ggeffects)
library(purrr)
library(dplyr)

# Grab original scales ONLY for back-transforming the Y axis
y_mean <- mean(model_data_scaled$mean_target_looking_critical_window, na.rm = TRUE)
y_sd   <- sd(model_data_scaled$mean_target_looking_critical_window, na.rm = TRUE)

pred_scaled <- map_dfr(sims, function(s) {
  res <- ggpredict(mods_scaled[[s]], terms = sprintf("scaled_%s [all]", s))
  
  res |> 
    as.data.frame() |> 
    transmute(
      sim_type   = s,
      similarity = x,                            # Keep X as the Z-score (-2 to +2 range)
      y          = (predicted * y_sd) + y_mean,  # Un-scale Y to 0-1
      lower      = (conf.low * y_sd) + y_mean,   # Un-scale CI lower
      upper      = (conf.high * y_sd) + y_mean   # Un-scale CI upper
    )
}) |> 
  mutate(sim_type = factor(sim_type, levels = sims))
```

## scaled?
```{r}
# Note: replace "scaled_similarity" with whatever name your pre-scaled x-variable 
# column has in your main 'plot_data' data frame.
clip_data_summarized_scaled <- summarize_similarity_data_collapsed_scaled(model_data_scaled, extra_fields = c("dataset_name"))
clip_data_summarized_scaled_bucket <- summarize_similarity_data_collapsed_scaled(model_data_scaled, extra_fields = c("dataset_name", "age_bucket"))

plot_data_scaled <- clip_data_summarized_scaled |>
    rename_with(~ gsub("scaled_", "", .x)) |>
  pivot_longer(
    cols = all_of(sims),
    names_to = "sim_type",
    values_to = "similarity"
  ) |>
  mutate(
    sim_type = factor(sim_type, levels = sims)
  )


fig <- ggplot(plot_data_scaled, aes(similarity, mean_value, 
                             colour = dataset_name, fill = dataset_name)) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  geom_point(size = 3, alpha = 0.1) +
  stat_smooth(method = "glm", geom = "line", alpha = 0.5, linewidth = 1) +
  
  # lmer population line + CI
  geom_ribbon(data = pred_scaled, inherit.aes = FALSE,
              aes(x = similarity, ymin = lower, ymax = upper),
              fill = "black", alpha = 0.15) +
  geom_line(data = pred, inherit.aes = FALSE,
            aes(x = similarity, y = y),
            colour = "black", alpha = 0.7, linewidth = 1.2) +
  
  facet_wrap(~ sim_type, nrow = 2, scales = "free_x",                        
             labeller = labeller(sim_type = facet_labs),
             strip.position = "top") +         
  scale_y_continuous(breaks = seq(0, 1, 0.2)) +
  coord_cartesian(ylim = c(0, 1)) +
  scale_color_discrete_qualitative(palette="Set 2") +
  labs(x = "Target–distractor similarity (z-scored)", 
       y = "Proportion target looking in critical window",
       colour = "Dataset Name", fill = "Dataset Name") +
  theme_classic(base_size=16) +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 20),
        panel.spacing = unit(1, "lines"),
        plot.margin = margin(t = 5.5, r = 30, b = 5.5, l = 5.5, unit = "pt"))
ggsave(here("figures", "dataset_similarities_zscored.png"),
       fig, width = 16, height = 13, bg = "white")
ggsave(here("figures", "dataset_similarities_zscored.svg"),
       fig, width = 16, height = 13, device="pdf")
```


## age
todo: ensure models used are not singular.
```{r}
model_data_age <- model_data_scaled |> 
  filter(!is.na(age_bucket))

### 2. Fit Individual Models using Scaled Data
mods_age <- expand_grid(
  sim = sims,
  age_bucket = unique(model_data_age$age_bucket)
) |>
  mutate(
    model = map2(sim, age_bucket, ~
      fit_main(
        sim = .x,
        data = model_data_age |> filter(age_bucket == .y)
      )
    ),
    name = paste(sim, age_bucket, sep = "__")
  )

# Convert into a named list of models
mods_age <- setNames(mods_age$model, mods_age$name)

# Note: Ensure 'model_data_age' holds the unscaled target variable, 
# or use your original unscaled base data frame here.
y_mean <- mean(model_data_age$mean_target_looking_critical_window, na.rm = TRUE)
y_sd   <- sd(model_data_age$mean_target_looking_critical_window, na.rm = TRUE)


# 2. Re-build the prediction grid, handling scaled predictors and un-scaling outcomes
pred <- map_dfr(names(mods_age), function(nm) {
  m     <- mods_age[[nm]]
  parts <- strsplit(nm, "__")[[1]]
  st    <- parts[1]   # Similarity type (sim)
  ab    <- parts[2]   # Age bucket
  
  # Find the scaled similarity term used in this specific model
  sim_col <- grep("scaled_.*_similarity$", all.vars(formula(m)), value = TRUE)
  if(length(sim_col) == 0) {
    # Fallback if your scaled variables don't strictly follow that exact pattern
    sim_col <- grep("_similarity$", all.vars(formula(m)), value = TRUE)
  }
  
  # Generate predictions based on the z-scored term
  ggpredict(m, terms = paste0(sim_col, " [all]")) |>
    as.data.frame() |>
    transmute(
      sim_type   = st, 
      age_bucket = ab,
      similarity = x,                         # Keep X as the Z-score (-2 to +2 range)
      y          = (predicted * y_sd) + y_mean,  # Un-scale Y to 0-1
      lower      = (conf.low * y_sd) + y_mean,   # Un-scale CI lower
      upper      = (conf.high * y_sd) + y_mean   # Un-scale CI upper
    )
}) |> 
  mutate(sim_type = factor(sim_type, levels = sims))
```


```{r}
# Note: Ensure plot_data_age_scaled has its similarity values z-scored 
# but uses the uniform column name 'similarity' after pivoting.

plot_data_scaled_bucket <- clip_data_summarized_scaled_bucket |>
    rename_with(~ gsub("scaled_", "", .x)) |>
  pivot_longer(
    cols = all_of(sims),
    names_to = "sim_type",
    values_to = "similarity"
  ) |>
  mutate(
    sim_type = factor(sim_type, levels = sims)
  )

age_scaled_plot <- ggplot(
  plot_data_scaled_bucket |> filter(!is.na(age_bucket)),
  aes(similarity, mean_value, colour = age_bucket, fill = age_bucket)
) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  geom_point(alpha = 0.12, size = 2) +
  
  # lmer population ribbon + lines split/colored by age_bucket
  geom_ribbon(
    data = pred |> filter(!is.na(age_bucket)),
    aes(x = similarity, ymin = lower, ymax = upper, fill = age_bucket),
    inherit.aes = FALSE, alpha = 0.2, colour = NA
  ) +
  geom_line(
    data = pred |> filter(!is.na(age_bucket)),
    aes(x = similarity, y = y, colour = age_bucket),
    inherit.aes = FALSE, linewidth = 1.2
  ) +
  
  facet_wrap(
    ~ sim_type,
    nrow = 2,
    scales = "free_x",
    labeller = labeller(sim_type = facet_labs),
    strip.position = "top"
  ) +
  scale_y_continuous(breaks = seq(0, 1, 0.2)) +
  coord_cartesian(ylim = c(0, 1)) +
  scale_color_viridis_d(option = "mako", direction=-1, begin=0.15, end=0.7) +  
  scale_fill_viridis_d(option="mako", direction=-1, begin=0.15, end=0.7) +
  labs(
    x = "Target–distractor similarity (z-scored)",
    y = "Proportion target looking in critical window",
    colour = "Age bucket",
    fill = "Age bucket"
  ) +
  theme_classic(base_size = 16) +
  theme(
    legend.position = "bottom",
    strip.text = element_text(size = 20),
    panel.spacing = unit(1, "lines"),
    plot.margin = margin(t = 5.5, r = 30, b = 5.5, l = 5.5, unit = "pt")
  )

ggsave(here("figures", "age_similarities_zscored.png"),
       age_scaled_plot, width = 15, height = 13, bg = "white")
```